Wide-angle tomographic measurement of the intergalactic gas pressure through galaxy-tSZ cross-correlations

### 2019-09-13 notes/paper/paper.tex
NK edits

\begin{abstract}
    We cross-correlate maps of the thermal Sunyaev-Zeldovich (tSZ)
    Compton-$y$ parameter published by \planck\ with the projected
    distribution of galaxies in a set of low-redshift tomographic
    bins. We use the nearly full-sky 2MASS Photometric Redshift and
    WISE $\times$ SuperCOSMOS public
    catalogues, covering the redshift range $z\lesssim0.4$. Our
    measurements allow us to place constraints on the redshift
    dependence of the mass-observable relation for tSZ cluster count
    analysis in terms of the so-called `hydrostatic bias' parameter
    $1-\bH$. These resultsts can also be interpreted as measurements
    of the bias-weighted average gas
    pressure $\langle bP_e\rangle$ as a function of redshift, a
    quantity that can be related to the thermodynamics of gas inside
    haloes and used to constrain energy injection processes. We
    measure $1-\bH$ with $\sim13\%$ precision in 6 equispaced
    redshift bins, and find no evidence for a redshift-dependent mass
    bias parameter, in agreement with previous analyses. Our mean value of
    $1-\bH = 0.59\pm0.03$ is also in good agreement with that estimated by
    the joint analysis of \planck\ cluster counts and CMB
    anisotropies. Our measurements of $\langle bP_e\rangle$, at the
    level of $\sim10\%$ in each bin, are the most stringent
    constraints on the redshift dependence of this parameter to date,
    and agree well both with previous measurements and with
    theoretical expectations from shock-heating
    models. \todo{Something we do not discuss at all, but should, is
      the fact that all our constraints assume (and depend on) the
      value of $\sigma_8$ found by Planck18 primary CMB. Our
      measurement of $1-\bH$ will depend on whatever prior on
      $\sigma_8$ we use, and therefore it shouldn't be completely
      surprising that we find something that agrees with \planck. It
      is however a good consistency check that we do, and what is
      definitely new is the measurement of the redshift
      dependence. I'd add words about this in Sections
      \ref{sec:results} and \ref{sec:conclusion}.}
\end{abstract}

@@

  In this work we will focus on the cross-correlation between galaxy clustering data and maps of the tSZ Compton-$y$ parameter, making use of existing data from the Planck collaboration \citep{2016A&A...594A..22P} and a set of photometric galaxy surveys \citep{2014ApJS..210....9B,2016ApJS..225....5B}. The availability of redshift information will allow us to place constraints on the cosmic evolution of the thermal gas pressure and the thermal energy in haloes, as well as to examine any redshift dependence of the mass bias for tSZ cluster studies. This is a relevant topic given the current mild tensions between SZ cluster number counts and CMB primary anisotropies \citep{2016A&A...594A..24P,2016ApJ...832...95D,2019ApJ...878...55B,2019arXiv190407887Z}, which could be caused by the assumptions made to model the $y$-mass relation. Conversely, under the assumption that the relation between mass and gas pressure is well understood, the tSZ effect can be thought of as a mass tracer, which can be used to break degeneracies with the galaxy bias parameter -- we will however leave this analysis for future work.

@@

\section{Theory}\label{sec:theory}
  Our work focuses on the cross-correlation of the projected galaxy overdensity in consecutive redshift bins $\delta_g$ and maps of the tSZ Compton-$y$ parameter.

  Here $\delta_g$ is simply the overdensity in the number of galaxies integrated over a redshift bin:
  \begin{equation}
    \delta_g(\nv)=\int dz\,\phi_g(z)\,\Delta_g(\chi(z)\,\nv),
  \end{equation}
  where $\nv$ is a unit vector on the sphere, $\chi(z)$ is the comoving radial distance to redshift $z$ and $\phi_g(z)$ is the normalised galaxy redshift distribution in the bin. $\Delta_g({\bf x})=n_g({\bf x})/\bar{n}_g-1$ is the 3D galaxy overdensity, where $n_g$ is the galaxy number density.

@@

    \subsubsection{Galaxies and the halo occupation distribution}\label{sssec:theory.hm.hod}
      To model the galaxy overdensity $\Delta_g$ we will use a Halo Occupation Distribution (HOD) model \citep{2002ApJ...575..587B,2005ApJ...633..791Z,2013MNRAS.430..725V}, as prescribed by \cite{2011ApJ...736...59Z}. The HOD models the galaxy content of dark matter haloes as being made up of central and satellite galaxies. Centrals lie at the centre of the halo, while satellites are distributed according to a profile $u_s(r|M)$. Haloes can have zero or one central, and the mean number of centrals for a halo of mass $M$ is modelled as a smoothed step function
      \begin{equation}
        \langle N_c(M)\rangle=\frac{1}{2}\left[1+{\rm erf}\left(\frac{\log(M/M_{\rm min})}{\sigma_{\rm lnM}}\right)\right].
      \end{equation}
      We then assume that satellites can only be formed if a halo has a central and has a mass larger than some threshold $M_0$. In that case, the average number of satellites follows a power law of the form:
      \begin{equation}
        \langle N_s(M)\rangle=N_c(M)\,\Theta(M-M_0)\,\left(\frac{M-M_0}{M_1'}\right)^{\alpha_s}.
      \end{equation}
      For simplicity, we will fix $M_0$ to $M_{\rm min}$, $\sigma_{\rm lnM}=0.15$ and $\alpha_s$=1, as in \cite{2018MNRAS.473.4318A}, leaving only two free parameters: $M_{\rm min}$ and $M_1'$.

@@

    \subsubsection{tSZ and pressure profiles}\label{sssec:theory.hm.pe}
      In order to describe the electron pressure in a halo, we use the generalised NFW profile (GNFW) described in \cite{2010A&A...517A..92A} and used in the \planck\ tSZ cluster analysis \cite{2016A&A...594A..24P}. This profile takes the form:
      \begin{equation}
        P_e(r)=P_*\,p(r/r_{500c}),
      \end{equation}
      where $r_{500c}$ is the cluster radius enclosing an overdensity of 500 times the critical density (see Section \ref{sssec:theory.hm.cm}). The normalization $P_*$ is given by
      \begin{equation}\label{eq:arnaud_norm}
        P_*=6.41\,\left(1.65\,\, {\rm eV}\,{\rm cm}^{-3}\right)h_{70}^{8/3}
        \left(\frac{h_{70}(1-\bH)M_{500c}}{3\times10^{14}M_\odot}\right)^{0.79},
      \end{equation}
      where $h_{70}=H_0/(70\ {\rm km}\,{\rm s}^{-1}\,{\rm Mpc}^{-1})$, and $M_{500c}$ is the halo mass enclosed by $r_{500c}$. The GNFW form factor is
      \begin{equation}
        p(x)=(c_P x)^{-\gamma}\left[1+(c_P x)^\alpha\right]^{(\gamma-\beta)/\alpha},
      \end{equation}
      with $(\alpha,\beta,\gamma,c_P)=(1.33,4.13,0.31,1.81)$. We must note that other pressure profiles have been proposed in the literature (e.g. \citejap{2012ApJ...758...75B}), but we choose this parameterisation in order to be able to relate our measurement of $(1-\bH)$ to the results of \cite{2016A&A...594A..24P}.

      The quantity $1-\bH$ in Eq.\!~\ref{eq:arnaud_norm} parameterises our lack of knowledge about the precise relation between mass and pressure in clusters. This factor is also commonly referred to as the `hydrostatic bias', since it was originally defined to account for the fraction of halo mass not in hydrostatic equilibrium missed by X-ray observations. Since this parameter also encapsulates other sources of bias in the X-ray-based mass estimates, we will instead refer to it as the \emph{mass bias}. Numerical simulations have constrained the mass deficit to be around 20\% (i.e. $1-\bH\simeq0.8$: \citejap{2012ApJ...758...74B}; \citejap{2014ApJ...782..107N}), although it is known that a smaller value is needed in order to fully reconcile CMB primary constraints and SZ cluster counts \citep{2016A&A...594A..24P}. This will be a central point of our discussion in Sections \ref{sec:results} and \ref{sec:conclusion}.

      Within the halo model description, and assuming a log-normal $y$-mass relation, the pressure profile cumulants are given by:
      \begin{align}
        &\langle u_y(k|M)\rangle=P_e(k),\\
        &\langle u_y^2(k|M)\rangle=P_e^2(k)\,e^{\sigma_{\ln Y}^2},
      \end{align}
      where $P_e(k)$ is the Fourier transform of the GNFW profile and $\sigma_{\ln Y}=0.173\pm0.023$ is the intrinsic logarithmic scatter in the $y$-mass relation \cite{2016A&A...594A..24P}. Note that, since we do not make use of the tSZ auto-correlation, we will not use the second order cumulant of the pressure profile in our analysis. However, since we analyse the galaxy-tSZ correlation, we need to model the covariance between the galaxy overdensity and pressure profiles. For simplicity, we adopt a simple parameterisation:
      \begin{equation}
        \langle u_y(k|M) u_g(k|M)\rangle = (1+\rho_{yg})\langle u_g(k|M)\rangle \langle u_y(k|M)\rangle,
      \end{equation}
      where $\rho_{yg}$ is a free parameter that determines the sign of the correlation between galaxy abundance and pressure. Marginalising over this parameter has the added advantage of removing any sensitivity of our final constraints to $1-\bH$ on the details of the cross-spectrum model in the 1-halo regime, where both parameters are completely degenerate.

      It is also worth exploring the halo model bias (Eq.\!~\ref{eq:hm_bias}) for the Compton-$y$ parameter. At $k\rightarrow0$ it is given by:
      \begin{align}\nonumber
        \langle bP_e\rangle&=\int dM\,\frac{dn}{dM}\,b_h(M)\,\int_0^\infty dr\,4\pi r^2\,P_e(r|M)\\\label{eq:by}
               &=\int dM\,\frac{dn}{dM}\,b_h(M)\,E_T(M),
      \end{align}
      where $E_T(M)$ is the thermal energy in a halo of mass $M$ \citep{2017MNRAS.467.2315V,2019arXiv190413347P}. A measurement of $\langle bP_e\rangle$ can therefore be related to the thermodynamics of gas inside haloes \MB{[more about this relation? and, this is electron pressure - how it connects to gas pressure?]} \DA{[Eq. \ref{eq:by} is the equation linking to thermal energy. Note that electron and total pressure are just proportional to each other, I've now noted this below Eq. \ref{eq:compty}.]}. We will also express our measurements in terms of this parameter in Section \ref{sec:results}.

    \subsubsection{Concentration-mass relation and mass definitions}\label{sssec:theory.hm.cm}
      Halo radii $r_\Delta$ are usually defined as the size of the sphere containing a given mass $M_\Delta$:
      \begin{equation}
        M_\Delta = \frac{4\pi}{3}\rho_*(z) \; \Delta \; r^3_\Delta.
      \end{equation}
      Common choices for $\rho_*$ are the critical density $\rho_c=3H^2(z)/4\pi G$ or the matter density $\rho_M(z)$ \MB{[which one do we use?]}\DA{[We state this in the next paragraph.]}. The spherical overdensity parameter $\Delta$ is usually chosen within the range $\sim(200,500)$ and sometimes defined as the quantity yielding the virial radius in the spherical top-hat collapse model $\Delta_v$ \citep{1998ApJ...495...80B}.

@@

\section{Data}\label{sec:data}
  \subsection{The Compton-$y$ map}\label{ssec:data.y}
    \begin{figure}
      \centering
      \includegraphics[width=0.5\textwidth]{mask_y.pdf}
      \includegraphics[width=0.5\textwidth]{mask_g.pdf}
      \caption{{\sl Top panel:} sky mask used for the Compton-$y$ map, corresponding to a sky fraction $f_{\rm sky}=0.59$. {\sl Bottom panel:} mask used for the 2MPZ and \wisc~ galaxy surveys, corresponding to a sky fraction $f_{\rm sky}=0.68$. The product of both masks leaves a usable sky fraction $f_{\rm sky}\simeq0.58$.
      %The top panel shows, in turquoise, the regions masked additionally when removing sky patches around the position of the tSZ sources detected by \planck\ above 6$\sigma$.
      }
      \label{fig:msk}
    \end{figure}
    We make use of the maps of the Compton-$y$ parameter made public by the Planck collaboration \citep{2016A&A...594A..22P}. These maps were generated using different flavours of the Internal Linear Combination method \citep{2004ApJ...612..633E,2008arXiv0811.4277V}. In a simplified description, the ILC technique selects the linear combination of all frequency channels that preserves the spectrum of the source one wishes to map minimising the map-level variance. The refined versions of the ILC method used by \planck\ further optimise the linear weights on different scales and different regions of the map, and project out sources with known spectra that are likely to cause significant contamination. In particular, the Planck collaboration has released two $y$ maps, extracted using the MILCA \citep{2013A&A...558A.118H} and NILC \citep{2011MNRAS.410.2481R} variations of the ILC technique. Both methods deproject CMB contamination through its well-known spectrum, but differ on the methods used to calculate the optimal scale-dependent and spatially-varying linear weights.

@@

  \subsection{2MPZ and \wisc}\label{ssec:data.g1}
    \begin{figure}
      \centering
      \includegraphics[width=0.5\textwidth]{nzs.pdf}
      \caption{Fiducial redshift distributions of the different galaxy samples used in this analysis. See Table \ref{tab:z_bins} for further details.}
      \label{fig:dndz}
    \end{figure}
    \begin{table}
      \begin{center}
        \begin{tabular}{l|cccc}
          \hline
          Sample & $[z_{{\rm ph},i},z_{{\rm ph},f}]$ & $\bar{z}$ & $\bar{n}_g\,[{\rm deg}^{-2^{\phantom{2}}}]$ & $\ell_{\rm max}$\\[1ex]
          \hline
          2MPZ    & N.A.         & 0.07 &  25.5 &  280 \\
          \wisc-1 & $[0.1,0.15]$ & 0.13 & 106   &  540 \\
          \wisc-2 & $[0.15,0.2]$ & 0.18 & 126   &  745 \\
          \wisc-3 & $[0.2,0.25]$ & 0.23 & 136   &  945 \\
          \wisc-4 & $[0.25,0.3]$ & 0.27 & 118   & 1130 \\
          \wisc-5 & $[0.3,0.35]$ & 0.32 & 41    & 1310 \\
          \hline
        \end{tabular}
        \caption{Galaxy samples used in this analysis, corresponding to the full 2MPZ survey and five tomographic redshift bins of the \wisc~ survey. The second, third and fourth columns list the photometric redshift interval defining the sample, its mean redshift and its number density respectively. The largest multipole used in the analysis of each sample (corresponding to a comoving scale of $k_{\rm max}\sim 1\,{\rm Mpc}^{-1}$) is shown in the last column.}\label{tab:z_bins}
      \end{center}
    \end{table} 
    We make use of two low-redshift photometric redshift catalogues, the 2MASS Photometric Redshift dataset \citep[2MPZ,][]{2014ApJS..210....9B} and the WISE $\times$ SuperCOSMOS catalogue \citep[\wisc,][]{2016ApJS..225....5B}. Both samples were created by cross-matching full-sky imaging surveys, and photo-$z$s were subsequently computed for all the included sources. Their broad photometric coverage, together with adequate spectroscopic calibration data allowed for obtaining well-constrained photo-$z$s, with minimal mean bias, relatively low scatter and a small number of outliers.

    2MPZ was constructed by cross-matching the extended-source catalogue from the 2 Micron All-Sky Survey \citep[2MASS,][]{2006AJ....131.1163S,2000AJ....119.2498J} with the photographic plates of SuperCOSMOS \citep{2001MNRAS.326.1295H,2016MNRAS.462.2085P} and the photometry of the Wide-field Infrared Survey Explorer (WISE: \citejap{2010AJ....140.1868W}). After applying an apparent magnitude cut $K_s<13.9$ to achieve uniformity, 2MPZ includes over 940,000 sources observed in 8 bands: SuperCOSMOS's optical $(B,R,I)$, 2MASS's near-infrared $(J,H,K_s)$ and WISE's mid-infrared $(W1,W2)$. Photometric redshifts (photo-$z$s) were extracted using the neural network code ANNz \citep{2004PASP..116..345C} trained on a large spectroscopic sample from overlapping surveys. The resulting photo-$z$'s have a typical error $\sigma_z\simeq0.015$, and the sample has a median redshift $z\simeq0.08$ (see also \citealt{2018MNRAS.476.1050B} for a more detailed analysis of 2MPZ photo-$z$ properties). As we will see, due to the low redshift of this sample, the cross-correlation with the $y$ map is dominated by the 1-halo term for 2MPZ, and little is gained by sub-dividing it into narrower redshift bins. Therefore we use 2MPZ as a single tomographic sample.

    A deeper sample is obtained by dropping the 2MASS data and cross-matching WISE and SuperCOSMOS only. After removing the sources already contained in 2MPZ, the resulting catalogue, \wisc, is $\sim3$ times deeper, contains $\sim20$ million sources and reaches up to redshift $z\sim0.4$, with a median redshift of $\sim0.2$. The photometric redshifts are less accurate, given the poorer photometric coverage $(B,\,R,\,W1,\,W2)$, with a mean error $\sigma_z/(1+z)\simeq0.035$. We divide the \wisc~ sample into 5 redshift bins, corresponding to photo-$z$ intervals of equal width $\delta z_{\rm photo}=0.05$ in the range $0.1<z_{\rm photo}<0.35$. Details regarding each of these redshift bins are given in Table \ref{tab:z_bins}, and the corresponding redshift distributions are shown in Fig.\!~\ref{fig:dndz}. We estimate fiducial redshift distributions for each tomographic sample as described in \cite{2018MNRAS.481.1133P}, and we discuss the marginalisation over uncertainties in these distributions in Section \ref{sssec:methods.syst.photoz}.

@@

    Although the jackknife method is able to provide an estimate of the size of the power spectrum uncertainties in a model independent way, it also has some drawbacks. There are only so many jackknife regions that can be selected with a reasonable size, and therefore the estimated covariance is noisy at some level. \DA{Typically the number of independent realizations used to estimate a sample covariance matrix should be at least one order of magnitude larger than the size of the data vector (up to 50 elements in our case).} Furthermore, since the footprint associated with the removal of each jackknife region is different from the overall footprint, the method is also not able to recover the mode-coupling associated with the map geometry by construction. In order to verify and improve our estimate of the covariance matrix, we make use of a second analytical estimator.

>>>>

\begin{abstract}
    We cross-correlate maps of the thermal Sunyaev-Zeldovich (tSZ)
    Compton-$y$ parameter published by \planck\ with the projected
    distribution of galaxies in a set of low-redshift tomographic
    bins. We use the nearly full-sky 2MASS Photometric Redshift and
    WISE $\times$ SuperCOSMOS public
    catalogues, covering the redshift range $z\lesssim0.4$. Our
    measurements allow us to place constraints on the redshift
    dependence of the mass-observable relation for tSZ cluster count
    analysis in terms of the so-called `hydrostatic bias' parameter
    $1-\bH$. These results can also be interpreted as measurements
    of the bias-weighted average gas
    pressure $\langle bP_e\rangle$ as a function of redshift, a
    quantity that can be related to the thermodynamics of gas inside
    haloes and used to constrain energy injection processes. We
    measure $1-\bH$ with $\sim13\%$ precision in 6 equispaced
    redshift bins, and find no evidence for a redshift-dependent mass
    bias parameter, in agreement with previous analyses. Our mean value of
    $1-\bH = 0.59\pm0.03$ is also in good agreement with that estimated by
    the joint analysis of \planck\ cluster counts and CMB
    anisotropies. Our measurements of $\langle bP_e\rangle$, at the
    level of $\sim10\%$ in each bin, are the most stringent
    constraints on the redshift dependence of this parameter to date,
    and agree well both with previous measurements and with
    theoretical expectations from shock-heating
    models.
\end{abstract}

@@

  In this work we focus on the cross-correlation between galaxy clustering data and maps of the tSZ Compton-$y$ parameter, making use of existing data from the Planck collaboration \citep{2016A&A...594A..22P} and a set of photometric galaxy surveys \citep{2014ApJS..210....9B,2016ApJS..225....5B}. The availability of redshift information will allow us to place constraints on the cosmic evolution of the thermal gas pressure and the thermal energy in haloes, as well as to examine any redshift dependence of the mass bias for tSZ cluster studies. This is a relevant topic given the current mild tensions between SZ cluster number counts and CMB primary anisotropies \citep{2016A&A...594A..24P,2016ApJ...832...95D,2019ApJ...878...55B,2019arXiv190407887Z}, which could be caused by the assumptions made to model the $y$-mass relation. Conversely, under the assumption that the relation between mass and gas pressure is well understood, the tSZ effect can be thought of as a mass tracer, which can be used to break degeneracies with the galaxy bias parameter -- we however leave this analysis for future work.

@@

\section{Theory}\label{sec:theory}
  Our work focuses on the cross-correlation of the projected galaxy overdensity in consecutive redshift bins, $\delta_g$, and maps of the tSZ Compton-$y$ parameter.

  Here, $\delta_g$ is simply the overdensity in the number of galaxies integrated over a redshift bin:
  \begin{equation}
    \delta_g(\nv)=\int dz\,\phi_g(z)\,\Delta_g(\chi(z)\,\nv),
  \end{equation}
  where $\nv$ is a unit vector on the sphere, $\chi(z)$ is the comoving radial distance to redshift $z$ and $\phi_g(z)$ is the normalised galaxy redshift distribution in the bin. $\Delta_g({\bf x})=n_g({\bf x})/\bar{n}_g-1$ is the 3D galaxy overdensity, where $n_g$ is the galaxy number density.

@@

    \subsubsection{Galaxies and the halo occupation distribution}\label{sssec:theory.hm.hod}
      To model the galaxy overdensity, $\Delta_g$, we use a Halo Occupation Distribution (HOD) model \citep{2002ApJ...575..587B,2005ApJ...633..791Z,2013MNRAS.430..725V}, as prescribed by \cite{2011ApJ...736...59Z}. The HOD models the galaxy content of dark matter haloes as being made up of central and satellite galaxies. Centrals lie at the centre of the halo, while satellites are distributed according to a profile $u_s(r|M)$. Haloes can have zero or one central, and the mean number of centrals for a halo of mass $M$ is modelled as a smoothed step function
      \begin{equation}
        \langle N_c(M)\rangle=\frac{1}{2}\left[1+{\rm erf}\left(\frac{\log(M/M_{\rm min})}{\sigma_{\rm lnM}}\right)\right].
      \end{equation}
      We then assume that satellites can only be formed if a halo has a central and has a mass larger than some threshold $M_0$. In that case, the average number of satellites follows a power law of the form:
      \begin{equation}
        \langle N_s(M)\rangle=N_c(M)\,\Theta(M-M_0)\,\left(\frac{M-M_0}{M_1'}\right)^{\alpha_s}.
      \end{equation}
      For simplicity, we fix $M_0$ to $M_{\rm min}$, $\sigma_{\rm lnM}=0.15$ and $\alpha_s$=1, as in \cite{2018MNRAS.473.4318A}, leaving only two free parameters: $M_{\rm min}$ and $M_1'$.

@@

    \subsubsection{tSZ and pressure profiles}\label{sssec:theory.hm.pe}
      In order to describe the electron pressure in a halo, we use the generalised NFW profile (GNFW) described in \cite{2010A&A...517A..92A} and used in the \planck\ tSZ cluster analysis \cite{2016A&A...594A..24P}. This profile takes the form:
      \begin{equation}
        P_e(r)=P_*\,p(r/r_{500c}),
      \end{equation}
      where $r_{500c}$ is the cluster radius enclosing an overdensity of 500 times the critical density (see Section \ref{sssec:theory.hm.cm}). The normalisation $P_*$ is given by
      \begin{equation}\label{eq:arnaud_norm}
        P_*=6.41\,\left(1.65\,\, {\rm eV}\,{\rm cm}^{-3}\right)h_{70}^{8/3}
        \left(\frac{h_{70}(1-\bH)M_{500c}}{3\times10^{14}M_\odot}\right)^{0.79},
      \end{equation}
      where $h_{70}=H_0/(70\ {\rm km}\,{\rm s}^{-1}\,{\rm Mpc}^{-1})$, and $M_{500c}$ is the halo mass enclosed by $r_{500c}$. The GNFW form factor is
      \begin{equation}
        p(x)=(c_P x)^{-\gamma}\left[1+(c_P x)^\alpha\right]^{(\gamma-\beta)/\alpha},
      \end{equation}
      with $(\alpha,\beta,\gamma,c_P)=(1.33,4.13,0.31,1.81)$. We must note that other pressure profiles have been proposed in the literature (e.g. \citejap{2012ApJ...758...75B}), but we choose this parameterisation in order to be able to relate our measurement of $(1-\bH)$ to the results of \cite{2016A&A...594A..24P}.

      The quantity $1-\bH$ in Eq.\!~\ref{eq:arnaud_norm} parameterises our lack of knowledge about the precise relation between mass and pressure in clusters. This factor is also commonly referred to as the `hydrostatic bias', since it was originally defined to account for the fraction of halo mass not in hydrostatic equilibrium missed by X-ray observations. Since this parameter also encapsulates other sources of bias in the X-ray-based mass estimates, we instead refer to it as the \emph{mass bias}. Numerical simulations have constrained the mass deficit to be around 20\% (i.e. $1-\bH\simeq0.8$: \citejap{2012ApJ...758...74B}; \citejap{2014ApJ...782..107N}), although it is known that a smaller value is needed in order to fully reconcile CMB primary constraints and SZ cluster counts \citep{2016A&A...594A..24P}. This will be a central point of our discussion in Sections \ref{sec:results} and \ref{sec:conclusion}.

      Within the halo model description, and assuming a log-normal $y$-mass relation, the pressure profile cumulants are given by:
      \begin{align}
        &\langle u_y(k|M)\rangle=P_e(k),\\
        &\langle u_y^2(k|M)\rangle=P_e^2(k)\,e^{\sigma_{\ln Y}^2},
      \end{align}
      where $P_e(k)$ is the Fourier transform of the GNFW profile and $\sigma_{\ln Y}=0.173\pm0.023$ is the intrinsic logarithmic scatter in the $y$-mass relation \cite{2016A&A...594A..24P}. Note that, since we do not make use of the tSZ auto-correlation, we do not use the second order cumulant of the pressure profile in our analysis. However, since we analyse the galaxy-tSZ correlation, we need to model the covariance between the galaxy overdensity and pressure profiles. For simplicity, we adopt a simple parameterisation:
      \begin{equation}
        \langle u_y(k|M) u_g(k|M)\rangle = (1+\rho_{yg})\langle u_g(k|M)\rangle \langle u_y(k|M)\rangle,
      \end{equation}
      where $\rho_{yg}$ is a free parameter that determines the sign of the correlation between galaxy abundance and pressure. Marginalising over this parameter has the added advantage of removing any sensitivity of our final constraints to $1-\bH$ on the details of the cross-spectrum model in the 1-halo regime, where both parameters are completely degenerate.

      It is also worth exploring the halo model bias (Eq.\!~\ref{eq:hm_bias}) for the Compton-$y$ parameter. At $k\rightarrow0$ it is given by:
      \begin{align}\nonumber
        \langle bP_e\rangle&=\int dM\,\frac{dn}{dM}\,b_h(M)\,\int_0^\infty dr\,4\pi r^2\,P_e(r|M)\\\label{eq:by}
               &=\int dM\,\frac{dn}{dM}\,b_h(M)\,E_T(M),
      \end{align}
      where $E_T(M)$ is the thermal energy in a halo of mass $M$ \citep{2017MNRAS.467.2315V,2019arXiv190413347P}. A measurement of $\langle bP_e\rangle$ can therefore be related to the thermodynamics of gas inside haloes \MB{[more about this relation? and, this is electron pressure - how it connects to gas pressure?]} \DA{[Eq. \ref{eq:by} is the equation linking to thermal energy. Note that electron and total pressure are just proportional to each other, I've now noted this below Eq. \ref{eq:compty}.]}. We also express our measurements in terms of this parameter in Section \ref{sec:results}.

    \subsubsection{Concentration-mass relation and mass definitions}\label{sssec:theory.hm.cm}
      Halo radii $r_\Delta$ are usually defined as the size of the sphere containing a given mass $M_\Delta$:
      \begin{equation}
        M_\Delta = \frac{4\pi}{3}\rho_*(z) \; \Delta \; r^3_\Delta.
      \end{equation}
      Common choices for $\rho_*$ are the critical density, $\rho_c=3H^2(z)/4\pi G$, or the matter density, $\rho_M(z)$ \MB{[which one do we use?]}\DA{[We state this in the next paragraph.]}. The spherical overdensity parameter $\Delta$ is usually chosen within the range $\sim(200,500)$ and sometimes defined as the quantity yielding the virial radius in the spherical top-hat collapse model $\Delta_v$ \citep{1998ApJ...495...80B}.

@@

\section{Data}\label{sec:data}
  \subsection{The Compton-$y$ map}\label{ssec:data.y}
    \begin{figure}
      \centering
      \includegraphics[width=0.5\textwidth]{mask_y.pdf}
      \includegraphics[width=0.5\textwidth]{mask_g.pdf}
      \caption{{\sl Top:} sky mask used for the Compton-$y$ map, corresponding to a sky fraction $f_{\rm sky}=0.59$. {\sl Bottom:} mask used for the 2MPZ and \wisc~ galaxy surveys, corresponding to a sky fraction $f_{\rm sky}=0.68$. The product of both masks leaves a usable sky fraction $f_{\rm sky}\simeq0.58$.
      %The top panel shows, in turquoise, the regions masked additionally when removing sky patches around the position of the tSZ sources detected by \planck\ above 6$\sigma$.
      }
      \label{fig:msk}
    \end{figure}
    We make use of the maps of the Compton-$y$ parameter made public by the Planck collaboration \citep{2016A&A...594A..22P}. These maps were generated using different flavours of the Internal Linear Combination method \citep{2004ApJ...612..633E,2008arXiv0811.4277V}. In a simplified description, the ILC technique selects the linear combination of all frequency channels that preserves the spectrum of the source one wishes to map minimising the map-level variance. The refined versions of the ILC method used by \planck\ further optimise the linear weights on different scales and different regions of the map, and project out sources with known spectra that are likely to cause significant contamination. In particular, the Planck collaboration has released two $y$ maps, extracted using the MILCA \citep{2013A&A...558A.118H} and NILC \citep{2011MNRAS.410.2481R} variations of the ILC technique. Both methods deproject CMB contamination through its well-known spectrum, but differ on the methods used to calculate the optimal scale-dependent and spatially-varying linear weights.

@@

  \subsection{2MPZ and \wisc}\label{ssec:data.g1}
    \begin{figure}
      \centering
      \includegraphics[width=0.5\textwidth]{nzs.pdf}
      \caption{Fiducial redshift distributions of the different galaxy samples used in this analysis. See Table \ref{tab:z_bins} for further details.}
      \label{fig:dndz}
    \end{figure}
    \begin{table}
      \begin{center}
        \begin{tabular}{l|cccc}
          \hline
          Sample & $[z_{{\rm ph},i},z_{{\rm ph},f}]$ & $\bar{z}$ & $\bar{n}_g\,[{\rm deg}^{-2^{\phantom{2}}}]$ & $\ell_{\rm max}$\\[1ex]
          \hline
          2MPZ    & N.A.         & 0.07 &  25.5 &  280 \\
          \wisc-1 & $[0.1,0.15]$ & 0.13 & 106   &  540 \\
          \wisc-2 & $[0.15,0.2]$ & 0.18 & 126   &  745 \\
          \wisc-3 & $[0.2,0.25]$ & 0.23 & 136   &  945 \\
          \wisc-4 & $[0.25,0.3]$ & 0.27 & 118   & 1130 \\
          \wisc-5 & $[0.3,0.35]$ & 0.32 & 41    & 1310 \\
          \hline
        \end{tabular}
        \caption{Galaxy samples used in this analysis, corresponding to the full 2MPZ survey and five tomographic redshift bins of the \wisc~ survey. The second, third and fourth columns list the photometric redshift interval defining the sample, its mean redshift and its number density respectively. The largest multipole used in the analysis of each sample (corresponding to a comoving scale of $k_{\rm max}\sim 1\,{\rm Mpc}^{-1}$) is shown in the last column.}\label{tab:z_bins}
      \end{center}
    \end{table} 
    We make use of two low-redshift photometric redshift catalogues, the 2MASS Photometric Redshift data set \citep[2MPZ,][]{2014ApJS..210....9B} and the WISE $\times$ SuperCOSMOS catalogue \citep[\wisc,][]{2016ApJS..225....5B}. Both samples were created by cross-matching full-sky imaging surveys, and photo-$z$s were subsequently computed for all the included sources. Their broad photometric coverage, together with adequate spectroscopic calibration data allowed for obtaining well-constrained photo-$z$s, with minimal mean bias, relatively low scatter and a small number of outliers.

    2MPZ was constructed by cross-matching the extended-source catalogue from the 2 Micron All-Sky Survey \citep[2MASS,][]{2006AJ....131.1163S,2000AJ....119.2498J} with the photographic plates of SuperCOSMOS \citep{2001MNRAS.326.1295H,2016MNRAS.462.2085P} and the photometry of the Wide-field Infrared Survey Explorer (WISE: \citejap{2010AJ....140.1868W}). After applying an apparent magnitude cut $K_s<13.9$ to achieve uniformity, 2MPZ includes over 940,000 sources observed in 8 bands: SuperCOSMOS's optical $(B,R,I)$, 2MASS's near-infrared $(J,H,K_s)$ and WISE's mid-infrared $(W1,W2)$. Photometric redshifts (photo-$z$s) were extracted using the neural network code ANNz \citep{2004PASP..116..345C} trained on a large spectroscopic sample from overlapping surveys. The resulting photo-$z$'s have a typical error $\sigma_z\simeq0.015$, and the sample has a median redshift $z\simeq0.08$ (see also \citealt{2018MNRAS.476.1050B} for a more detailed analysis of 2MPZ photo-$z$ properties). As demonstrated in Section \ref{sec:results}, due to the low redshift of this sample, the cross-correlation with the $y$ map is dominated by the 1-halo term for 2MPZ, and little is gained by sub-dividing it into narrower redshift bins. Therefore we use 2MPZ as a single tomographic sample.

    A deeper sample is obtained by ignoring the 2MASS data and cross-matching WISE and SuperCOSMOS only. After removing the sources already contained in 2MPZ, the resulting catalogue, \wisc, is $\sim3$ times deeper, contains $\sim20$ million sources and reaches up to redshift $z\sim0.4$, with a median redshift of $\sim0.2$. The photometric redshifts are less accurate, given the poorer photometric coverage $(B,\,R,\,W1,\,W2)$, with a mean error $\sigma_z/(1+z)\simeq0.035$. We divide the \wisc~ sample into 5 redshift bins, corresponding to photo-$z$ intervals of equal width $\delta z_{\rm photo}=0.05$ in the range $0.1<z_{\rm photo}<0.35$. Details regarding each of these redshift bins are given in Table \ref{tab:z_bins}, and the corresponding redshift distributions are shown in Fig.\!~\ref{fig:dndz}. We estimate fiducial redshift distributions for each tomographic sample as described in \cite{2018MNRAS.481.1133P}, and we discuss the marginalisation over uncertainties in these distributions in Section \ref{sssec:methods.syst.photoz}.

@@

    Although the jackknife method is able to provide an estimate of the size of the power spectrum uncertainties in a model independent way, it also has some drawbacks. There are only so many jackknife regions that can be selected with a reasonable size, and therefore the estimated covariance is noisy at some level. \DA{Typically the number of independent realisations used to estimate a sample covariance matrix should be at least one order of magnitude larger than the size of the data vector (up to 50 elements in our case).} Furthermore, since the footprint associated with the removal of each jackknife region is different from the overall footprint, the method is also not able to recover the mode-coupling associated with the map geometry by construction. In order to verify and improve our estimate of the covariance matrix, we make use of a second analytical estimator.
